---
title: "Circular GLMMs for bat vocalizations"
output:
  html_document:
    df_print: paged
---


###Background
We conducted an experiment in which we recorded vocalizations from groups of freely
socializing bats. During acoustic recordings, we played three different kinds of 
white noise. This white noise was amplitude modulated either at 8Hz or 15Hz. We 
wanted to know whether the bats preferred to vocalize at certain points in the 
ongoing amplitude modulation cycle for each of the different types of noise. 

###Aim
Using circular GLMMs, we would like to model and describe the effect of 
**white noise condition** x **amplitude modulation rate** on the **call onsets**. 

Call onsets are provided in the sample data in two ways, for each call event (row)
we have recorded the call onset in seconds (`start.per`) and in radians (`start.phase`) relative
to the start of the amplitude modulation cycle in which the call was emitted. 

We will use the call onsets in radians for analysis because the two modulation rate
conditions have different periods (125ms and 67ms, respectively), so using call timing
within the cycle in radians allows us to compare across these two conditions:

###Data

Outcome variable:
*`start.phase`: call onsets in radians, 0-2pi; 0/2pi represents the amplitude trough, 
while pi represents the amplitude peak

Predictor variables:
*`condition`: white noise types, either "A", "B", or "C"
*`modulation`: rate at which the white noise was amplitude modulated, either 8 or 15Hz

Grouping variables are:
*`group`: we tested four groups of bats, labelled 1-4; this should be used as a grouping/
random effects variable
*`session`: each group was tested on 5 days, 1-5; this can be used to determine 
whether changes in calling patterns emerged over time
*`minute`: this indicates the minute since the start of the current recording
condition when the call occurred; each condition was recorded for ~1 hour; note that 
some rows have NA's here; can also be used to track emergence of call patterns

###Model
Instantaneous phase of call onsets predicted by white noise type and amplitude modulation
rate, for each group of animals.

$\theta_{call} ~ white noise * modulation + group$


#### Setup
```{r Setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(circular)
library(bpnreg)
set.seed(42)
```

Load in our sample data. 
```{r Load}
calls <- read.csv("sample_data.csv")
```

Convert start.phase to circular data (is this necessary?)...
```{r Recode}
glm.calls <- calls %>% mutate(phase.circ = circular(start.phase, 
                                                 units = "radians", 
                                                 zero = 0, 
                                                 modulo = "asis", 
                                                 rotation = "counter"), .after = start.phase) %>%
                      # dummy coding
                      mutate(condition = as.factor(as.numeric(as.factor(condition))-1), 
                             modulation = as.factor(as.numeric(as.factor(modulation))-1),
                             group = as.numeric(group)) %>%
                      select(-minute)

head(glm.calls)

# double check range of outcome variable (in radians, 0-2pi)
min(glm.calls$phase.circ)
max(glm.calls$phase.circ)
```

#### Models

##### Fixed-effects only
Circular coefficients provide summary statistics (mean, mode, standard deviation 
and 95% highest posterior density interval) of the posterior samples:
```{r}
# just condition
m1 <- bpnr(phase.circ ~ condition,
             data = glm.calls, its = 1000, burn = 100, n.lag = 3, seed = 42)

  coef_circ(m1, type = "categorical", units = "radians")

# condition + modulation  
m2 <- bpnr(phase.circ ~ condition + modulation,
             data = glm.calls, its = 1000, burn = 100, n.lag = 3, seed = 42)

  coef_circ(m2, type = "categorical", units = "radians")

# condition * modulation
m2prime <- bpnr(phase.circ ~ condition * modulation,
             data = glm.calls, its = 1000, burn = 100, n.lag = 3, seed = 42)

  coef_circ(m2prime, type = "categorical", units = "radians")
```

Can also model the effect of condition for modulation rates separately:
```{r eval=FALSE, include=FALSE}
# model with modulation = 8Hz only
m1.8 <- bpnr(phase.circ ~ condition,
             data = glm.calls[glm.calls$modulation==0,], its = 100, burn = 50, n.lag = 3, seed = 42)
  
  coef_circ(m1.8, type = "categorical", units = "radians")

# model with modulation = 15Hz only
m1.15 <- bpnr(phase.circ ~ condition,
             data = glm.calls[glm.calls$modulation==1,], its = 100, burn = 50, n.lag = 3, seed = 42)

  coef_circ(m1.15, type = "categorical", units = "radians")
```


##### Fixed and random effects

Note that this runs only on our cluster and not on my local computer (32 GB RAM + a GPU). 
I'd appreciate any tips for optimizing that may allow it to run on a computer with 
that capacity. :)
```{r eval=FALSE}
mm0 <- bpnme(phase.circ ~ (1|group), 
             data = glm.calls, its = 100, burn = 100, n.lag = 3, seed = 42)

mm1 <- bpnme(phase.circ ~ conditon + (1|group), 
             data = glm.calls, its = 1000, burn = 100, n.lag = 3, seed = 42)

mm2 <- bpnme(phase.circ ~ condition + modulation + (1|group), 
             data = glm.calls, its = 1000, burn = 100, n.lag = 3, seed = 42)

mm2prime <- bpnme(phase.circ ~ condition * modulation + (1|group), 
             data = glm.calls, its = 1000, burn = 100, n.lag = 3, seed = 42)
```

## Issues/Questions

1. While my outcome variable (in radians) consists of values between 0 and 2pi, I
get posterior means that are negative. I can understand why posterior mean *differences*
may be negative, but am somewhat confused by the results here. 

2. One of the main comparisons of interest for our study is the difference in
the distribution of the outcome variable for `condition`==0 ("A") and `condition`==1 ("B"),
conditional on the value of `modulation`.  
This means that the posterior means are readily usable for us. However, I would also 
like to obtain Bayes Factors for the differences in the circular means between 
the intercept and the two other levels. However, the following does not seem to be possible:

```{r}
#BFc(m2prime, hypothesis = "(Intercept) < condition2")
```
How might I go about this, shy of running a different model where I switch the coding
of the conditions?

3. Regression curves: I would like to obtain something like a regression curve that I could plot over
existing data. However, it seems that I cannot get regression curve parameters
without a continuous predictor.
Is there a structural (mathematical) reason for this, or has it perhaps just not 
yet been implemented?

4. Prediction: I would like to use the model to generate predicted values. However, 
trying to do this by passing the model object to R's `predict()` returns an error: 
`no applicable method for 'predict' applied to an object of class "c('bpnr', 'list')"`...

```{r}
#predict(m2prime)
```

Is there a native way to extract predicted values from the circular models? If not, 
can you recommend a workaround?
Relatedly, I was not sure if this is what is obtained by doing the following: 
```{r eval=FALSE, include=FALSE}
#head(m2prime$beta1)
#head(m2prime$beta2)
```


5. Visualizing model outcomes: With categorical predictors, are there any other
estimated/predicted parameters that one can visualize other than predicted means?
For example, confidence intervals or dispersion/concentration parameters (equivalent to
Rao's test for dispersion)?

## Feature Requests
- Would it be possible to display variances into the model output? Similar to how lme4
provides info on variance explained by each predictor?


### Notes for self from the documentation

If predictors are continuous:
* `ax` = the location of the inflection point of the regression curve on the axis of the predictor.
* `ac` = the location of the inflection point of the regression curve on the axis of the circular outcome.
* `bc` = the slope of the tangent line at the inflection point. An increase of 1 unit of the predictor at the inflection point leads to a `bc` change in the circular outcome.
* `AS` = the average slopes of the circular regression. An increase of 1 unit of the predictor leads to a `AS` change in the circular outcome on average.
* `SAM` = the circular regression slopes at the mean. An increase of 1 unit of the predictor leads to a `SAM` change in the circular outcome at the average predictor value.
* `SSDO`= the signed shortest distance to the origin.

If predictors are categorical:
The output returns summary statistics for the posterior distributions of the circular means for all categories and combination of categories of the categorical variables in the model, as well as differences between these means. 


Fit statistics:
All five fit statistics are computed as in Gelman et.al. (2014). The `lppd` is
an estimate of the expected log predictive density, the `DIC` is the Deviance
Information Criterion, the `DIC_alt` is a version of the DIC that uses a
slightly different definition of the effective number of parameters, the `WAIC1`
and `WAIC2` are the two versions of the Watanabe-Akaike or Widely Available
Information Criterion presented in Gelman et.al. (2014).

Raw posterior estimates are stored in the following objects:

* `a.x` = posterior samples for the the locations of the inflection point of the regression curve on the axis of the predictor.
* `a.c` = posterior samples for the the locations of the inflection point of the regression curve on the axis of the circular outcome.
* `b.c` = posterior samples for the slopes of the tangent line at the inflection point. 
* `AS` = posterior samples for the average slopes of the circular regression.
* `SAM` = posterior samples for the circular regression slopes at the mean.
* `SSDO`= posterior samples for the signed shortest distance to the origin.
* `circ.diff` = posterior samples for the circular differences between intercept and other categories of categorical variables.
* `beta1` = posterior samples for the fixed effects coefficients for the first component.
* `beta2` = posterior samples for the fixed effects coefficients for the second component.

In circular mixed-effects models the following additional parameters can be obtained:

* `b1` = posterior samples for the random effects coefficients for the first component.
* `b2` = posterior samples for the random effects coefficients for the second component.
* `circular.ri` = posterior samples for the circular random intercepts for each individual.
* `omega1` = posterior samples for the random effect variances of the first component.
* `omega2` = posterior samples for the random effect variances of the first component.
* `cRS` = posterior samples for the circular random slope variance.
* `cRI` = posterior samples of the mean resultant length of the circular random intercept, a measure of concentration. 



